Setup_OMs

### 2023-09-10 Ecov_study/catchability/Catchability_ecov_sims.Rmd
Add code updates, separate functions into separate scripts.

---
title: "Catchability_ecov_sims"
output: html_document
date: "2022-11-14"
---
# Function definitions
Functions based on simulation testing vignette: https://github.com/timjmiller/wham/blob/devel/vignettes/ex10_simulation.Rmd

## Generate simulated OM data
```{r}
#' @param OM A fitted WHAM model to use as the operating model, no default.
#' 
#' @return A simulated OM dataset, may be used to overwrite input$data either entirely or in part in simulations

simOM <- function(OM=NULL){
        input <- OM$input
        dataOM = OM$simulate(complete=TRUE) # Overwrite input with simulated data & return
        return(dataOM)
}
```

## Fit specified EMs to simulated OM data and store results
```{r}
#' @param nsim A number specifying how many simulations to run, default = 1.
#' @param OM A fitted WHAM model to use as an operating model, no default.
#' @param inputEMlist A list of WHAM input data objects for all estimation models, no default.
#' @param outdir A filename specifying the directory where time-stamped results should be stored, if NULL (default) then the results are not saved as RData,
#' 
#' @return A nested list with names and results corresponding to the operating models and all estimation models. For each model results contains a sub-list with items for each simulation's results stored as sub-sub lists (because WHAM results are stored as a list). If outdir != NULL then results are saved as RData and returned as an object from the function.

simTestWHAM <- function(nsim = 1,
                        OM = NULL,
                        inputEMlist = NULL,
                        outdir = NULL){
        # Simulation set-up
        obs_names = c("agg_indices","agg_catch","catch_paa","index_paa", "Ecov_obs", "obsvec") # Data overwritten by OM
        namesEM <- lapply(inputEMlist, FUN = function(inputEMlist){inputEMlist$model_name}) %>% unlist() # EM model names

        # Storage set-up
        results <- NULL #!!!!! Still need to work on setting this up
        storeEM <- lapply(namesEM, FUN = assign, value = list())
        names(storeEM) <- namesEM

        # Generate OM data for all simulations
        sim_inputs <- replicate(nsim, simOM(OM), simplify = FALSE)

        for(isim in 1:nsim){ # Loop over simulations
                for(iEM in 1:length(inputEMlist)){ # Loop over estimation models
                        # Pull EM initial input from list
                        inputEM <- inputEMlist[[iEM]]

                        # Update EM input with OM simulated data
                        inputEM$data[obs_names] = sim_inputs[[isim]][obs_names]

                        # Fit WHAM model to updated EM input and save results
                        resultEM <- fit_wham(inputEM, do.osa = FALSE, do.retro = FALSE, MakeADFun.silent = TRUE) #!!!! Figure out how to store

                        # Store model-specific results for given isim here (append EM results to storage list for that EM)
                        storeEM[which(namesEM == namesEM[iEM])][[1]] <- append(storeEM[which(namesEM == namesEM[iEM])], list(resultEM))

                } # End loop over EM
        } # End loop over nsims

        # Store all results
        results$OM <- sim_inputs
        names(results) <- OM$model_name # Name OM column based on model_name
        results <- append(results, storeEM) # Should add named EM results to the final resuts object

        # Save time-stamped results as RData if output directory provided as argument
        timeStamp <- Sys.time() %>% gsub(" ", "_",.) %>% gsub(":", "-", .) # Change millisecond half of Sys.time() output to avoid having spaces/weird characters in filenames
        if(is.null(outdir)==FALSE){
                save(results, file = paste(outdir, paste0("simWHAM_", nsim, "_nsim_", OM$model_name, "_OM_", timeStamp, ".RData"), sep="/"))
        }

        return(results) # Always return results as an object
}
```

### Visual describing results structure
WHAM results objects only displayed for OM simulation 1, similar structure for other simulations and for estimation models.
```{r}
library(DiagrammeR)

@@

## Post-process results
```{r}
#' @param filenames A vector of strings (including .Rda extensions) indicating what Rdata results files to read in and post process
#' 
#' @return

postprocess_simTestWHAM <- function(filenames = NULL){
        # Set up storage for processed results

        # Loop over files
        for(ifile in 1:length(filenames)){
                # Read in Rdata object
                fileDat <- readRDS(filenames[ifile])

                nsim <- # Based on length of results in given filename

                # Loop over simulations in each file
                for(isim in 1:nsim){
                        # Calculate performance metrics here & add to storage
                }

        } # End loop over filenames
}
```

# Specifics for environmentally-linked catchability WHAM runs
## Define Operating Models (OMs)
All OMs should have catchability linked to an environmental covariate but may differ in 
- Uncertainty around this relationship #!!!!!!!!
- Sign of this relationship (+/-) #!!!!!!!!

Set up input and fit OM1
```{r}
inputOM1 <- prepare_wham_input(model_name = "OM1") #!!!!!!!!!!!!
OM1 <- fit_wham(inputOM1, do.osa = FALSE, do.retro = FALSE, MakeADFun.silent = TRUE)
```

## Define Estimation Models (EMs)
Correctly specified link between catchability and the environment
```{r}
EMq_env <- prepare_wham_input(model_name = "EMq_env")
```

Catchability random effect and a link between catchability and the environment
```{r}
EMq_rand_env <- prepare_wham_input(model_name = "EMq_rand_env")
```

Only a catchability random effect, fit to environmental model without link
```{r}
EMq_rand <- prepare_wham_input(model_name ="EMq_rand")
```

No catchability considerations (i.e. status quo), fit to environmental model without link !!!! how does this work if you feed in the ecov_beta?
```{r}
EMq <- prepare_wham_input(model_name = "EMq")
```

## Simulation testing 
```{r}
simTestWHAM(nsim = 5, OM = OM1, inputEMlist = list(EMq_env, EMq_rand_env, EMq_rand, EMq)) # Update once OM/EM specifics finalized!!!!

@@

## Post-process simulation results
```{r}
postprocess_simTestWHAM(filenames = c("filename1.Rds", "filename2.Rds", "ect.Rds")) # !!! Update once completed runs
```

!!!! post-process based on storage & how sims run
- could run sims based varying isim
- could also loop over OM but this could be computationally expensive
- I recommend labeling runs with OM and EM names (and setting these up with informative) so it is easy to pull out results & compare based on model setup
  - pull from mod$model_name to help aggregate data

@@

### Result formatted as list of lists, save as rdata object in specified directory with timestamp in name
run1_11_14_22.Rda<-result
 - EM1 
 - EM2 
 - EM3
 - EM4
 - OM result

# Other things to consider
- Other EMs can be developed (e.g. with alternative ecov process assumptions) but should be given DIFFERENT model_names to help with post-processing

# Short script for function testing purposes only!!!
```{r}
library(wham)
# For testing function
testdata <- read_asap3_dat(paste("/Users/amandahart/Research/Plaice_Methods_Development", "data", "PlaiceWHAM-2019_revised_NEFSC-LW-WAA_splitNEFSC-BigUnit_fixSeason.DAT", sep="/")) # this can be swapped out for any ASAP file you have to set up a WHAM model for debugging
testinput <- prepare_wham_input(testdata)
testWHAM <- fit_wham(testinput, MakeADFun.silent = TRUE, do.osa = FALSE, do.retro=FALSE)

@@

>>>>

---
title: "Catchability_ecov_sims"
output: html_document
date: "2022-11-14"
editor_options: 
  markdown: 
    wrap: 72
---

Load packages & source functions used in simulation testing
```{r}
# Packages
library(tidyverse)
library(wham)
# Simulation testing functions
source(here::here("Ecov_study","catchability","simOM.R"))
source(here::here("Ecov_study","catchability","simTestWHAM.R"))
```

### Visual describing results structure

WHAM results objects only displayed for OM simulation 1, similar
structure for other simulations and for estimation models.

```{r}
library(DiagrammeR)

@@

## Post-process results

```{r}

@@

# Specifics for environmentally-linked catchability WHAM runs

## Generic stock info that is never misspecified/changed
```{r}
# Pull generic groundfish life history/data
source(file.path(here::here(),"common_code", "make_basic_info.R"))
groundfish_info <- make_basic_info()

# Assume logistic selectivity for the fleet and all indices
sel_list <- list(model = c(rep("logistic", groundfish_info$n_fleets),rep("logistic", groundfish_info$n_indices)),
                   initial_pars = rep(list(c(5,1)), groundfish_info$n_fleets + groundfish_info$n_indices),
                 n_selblocks = c(groundfish_info$n_fleets + groundfish_info$n_indices)) #fleet, index

# Starting mean M estimate at 0.2, estimate constant M for all ages/years
M <- list(initial_means = rep(0.2, length(groundfish_info$ages)))

# Assume random about the mean recruitment, beverton-holt stock recruit didn't converge with an invertible hessian
recruit_model <- 2 
# NAA_list <- list(sigma="rec") # Recruitment random effects so stock recruit function implemented

genericInput <- prepare_wham_input(basic_info = groundfish_info, 
                                   selectivity = sel_list, 
                                   recruit_model = recruit_model) #,
                                   #NAA_re = NAA_list)
genericFit <- fit_wham(genericInput, MakeADFun.silent = TRUE, do.retro = FALSE, do.osa = FALSE)

## Define Operating Models (OMs)

All OMs should have catchability linked to an environmental covariate
but may differ in - Uncertainty around this relationship #!!!!!!!! -
Sign of this relationship (+/-) #!!!!!!!!

Set up input and fit OM1

```{r}
inputOM1 <- prepare_wham_input(model_name = "OM1") #!!!!!!!!!!!!
OM1 <- fit_wham(inputOM1, do.osa = FALSE, do.retro = FALSE, MakeADFun.silent = TRUE)
```

## Define Estimation Models (EMs)

| q-Env link                    | Seasonal misspecification |
| ----------------------------- | ------------------------- |
| Without link                  | Both seasons correct (q-Env link in EM and OM)         |
| With link                     | 1 season misspecified (q-Env linked but no link in OM) |
| Without link, q random effect | Both seasons misspecified (no link but link in OM)     |
| With link, q random effect    | |

Correctly specified link between catchability and the environment

```{r}
EMq_env <- prepare_wham_input(model_name = "EMq_env")
```

Catchability random effect and a link between catchability and the
environment

```{r}
EMq_rand_env <- prepare_wham_input(model_name = "EMq_rand_env")
```

Only a catchability random effect, fit to environmental model without
link

```{r}
EMq_rand <- prepare_wham_input(model_name ="EMq_rand")
```

No catchability considerations (i.e. status quo), fit to environmental
model without link !!!! how does this work if you feed in the ecov_beta?

```{r}
EMq <- prepare_wham_input(model_name = "EMq")
```

## Simulation testing

```{r}
simTestWHAM(nsim = 5, OM = OM1, inputEMlist = list(EMq_env, EMq_rand_env, EMq_rand, EMq)) # Update once OM/EM specifics finalized!!!!

## Post-process simulation results

```{r}
postprocess_simTestWHAM(filenames = c("filename1.Rds", "filename2.Rds", "ect.Rds")) # !!! Update once completed runs
```

!!!! post-process based on storage & how sims run - could run sims based
varying isim - could also loop over OM but this could be computationally
expensive - I recommend labeling runs with OM and EM names (and setting
these up with informative) so it is easy to pull out results & compare
based on model setup - pull from mod\$model_name to help aggregate data

@@

### Result formatted as list of lists, save as rdata object in specified directory with timestamp in name

run1_11_14_22.Rda\<-result - EM1 - EM2 - EM3 - EM4 - OM result

# Other things to consider

-   Other EMs can be developed (e.g. with alternative ecov process
    assumptions) but should be given DIFFERENT model_names to help with
    post-processing

# Short script for function testing purposes only!!!

```{r}
library(wham)
# For testing function
testdata <- read_asap3_dat(paste("/Users/amandahart/Research/Plaice_Methods_Development", "data", "PlaiceWHAM-2019_revised_NEFSC-LW-WAA_splitNEFSC-BigUnit_fixSeason.DAT", sep="/")) # this can be swapped out for any ASAP file you have to set up a WHAM model for debugging
testinput <- prepare_wham_input(testdata)
testWHAM <- fit_wham(testinput, MakeADFun.silent = TRUE, do.osa = FALSE, do.retro=FALSE)

@@

# To do next

- Finish OM setup
  - change how ecov passed in
  - set up range of uncertainty values to explore
  - tim sets equilibrium F (eq_F_init)
    - setup OM without fit & use model to find F40 proxy
  - Tim also sets intitial numbers at age based on equilibrium (overfished F=eqFN1*max_mult_Fmsy or not eqFN1=0)
- Tim lines 71-79 pf make_om use internal processing for prepare_wham_input to plug in setup from M_Ecov_om_setup.R
- Update what is saved as output
  - calculate performance metrics
- Run simulations, hpcc?
- Separate sims and function scripts

### 2023-09-10 Ecov_study/catchability/Setup_OMs.Rmd
Add code updates, separate functions into separate scripts.

-   Ecov ar1 process on catchability for both surveys (spring & fall)
    with mean = 0

-

@@

-   Fishing history = F_MSY for 40 year history OR fish 2.5\*F_MSY for
    20 years than fish at F_MSY for remaining 20

-

@@

-

```{r}
##### OM
# Ecov process & effect magnitude set up
Ecov_process_sig <- c(0.1, 0.5)
Ecov_process_cor <- c(0, 0.5)
Ecov_process_obs_sig <- c(0.1, 0.5)
Ecov_effect <- c(0, 0.25, 0.5) # beta

@@

# Observation error
ageComp_sig <- c(0.3, 1.5) # Fleet and index age comp treated with same error
logIndex_sig <- c(0.1, 0.4)

# Factorial combinations
OMsetup <- expand.grid(Eco_process_sig = Ecov_process_sig, Ecov_process_cor = Ecov_process_cor, Ecov_process_obs_sig = Ecov_process_obs_sig, Ecov_effect = Ecov_effect, F_hist = F_hist, ageComp_sig = ageComp_sig, logIndex_sig = logIndex_sig)
OMsetup

# Catchability set up
q_Ecov <- c("includeEcov", "excludeEcov") # May want to include these as a sensitivity run to compare what happens if OM has no ecov effect but EM assumes there is one (ecov process setup won't matter since fitting to data without effect specified in model)

##### OM sensitivity runs
# Factorial combinations when on ecov effect in OM
expand.grid(q_Ecov = q_Ecov, F_hist = F_hist, ageComp_sig = ageComp_sig, logIndex_sig = logIndex_sig) # probably do a subset, we have to decide what subset

##### EM
# EM factorial settings
miss_season <- c("BOTH", "ONE", "NONE") # Number of seasons with correctly specified effect, assumes 2 survey indices (spring & fall)
miss_q <- c("NoEnv", "NoEnvRand", "WithEnvRand", "WithEnv") # Catchability setup for EM
expand.grid(miss_season = miss_season, miss_q = miss_q)
```

-   maybe don't vary obs/process NAA since we are testing ability to
    recover catchability effect, could do sensitivity run with survival
    random effects and pick one of these settings

@@

# Loop over OMsetup to initialize OMs
for(iom in 1:nrow(OMsetup)){
  # Pull generic input settings
  input <- genericInput

  # Ecov setup
    Ecov <- list(label = "Ecov",
                 lag = 0, 
                 where = "q", # Where/how/indices settings need to change if we do sensitivity runs
                 how = 1,
                 indices = c(1,2),
                 process_model = "ar1", 
                 process_mean_vals = rep(0,length(groundfish_info$years)), # End generic inputs for ecov
                 process_sig_vals = rep(OMsetup$Eco_process_sig[iom]),
                 process_cor_vals = rep(OMsetup$Ecov_process_cor[iom]),
                 beta_vals = list(rep(list(matrix(rep(OMsetup$Ecov_effect[iom], length(groundfish_info$ages)), nrow = 1)), 4)), #!!! Why 4 rows?
                 logsigma = as.matrix(rep(OMsetup$Ecov_process_obs_sig[iom], length(groundfish_info$years))))

@@

  # 
  # Fishing history
F_hist <- c("H-MSY", "MSY") # High-then MSY vs. MSY for entire history

# Observation error
ageComp_sig <- c(0.3, 1.5) # Fleet and index age comp treated with same error
logIndex_sig <- c(0.1, 0.4)

>>>>

-   Ecov ar1 process on catchability with mean = 0

@@

-   Fishing history = F_MSY for 40 year history OR fish 2.5\*F_MSY for
    20 years than fish at F_MSY for remaining 20 - NOT CURRENTLY IMPLEMENTED, IN PROGRESS

-   Seasonality of Ecov effect on catchability (neither, one, or both seasons impacted) - NOT CURRENTLY IMPLEMENTED, IN PROGRESS

@@

-   Seasonality of Ecov effect on catchability

Run analysis in 3 pieces:
1) Factorial runs with different ecov process magnitude, fishing history and observation error
-   maybe don't vary obs/process NAA since we are testing ability to
    recover catchability effect, could do sensitivity run with survival
    random effects and pick one of these settings
```{r}
##### OM
# Ecov process & effect magnitude set up
Ecov_process_sig <- c(0.1, 0.5)
Ecov_process_cor <- c(0, 0.5)
Ecov_process_obs_sig <- c(0.1, 0.5)
Ecov_effect <- c(0, 0.25, 0.5) # beta

@@

# Observation error
ageComp_sig <- c(0.3, 1.5) # Fleet and index age comp treated with same error
log_index_sig <- c(0.1, 0.4) # agg_index_sigma
log_catch_sig <- 0.1 # agg_catch_sigma

# Factorial combinations
OMsetup <- expand.grid(Eco_process_sig = Ecov_process_sig, Ecov_process_cor = Ecov_process_cor, Ecov_process_obs_sig = Ecov_process_obs_sig, Ecov_effect = Ecov_effect, F_hist = F_hist, ageComp_sig = ageComp_sig, logIndex_sig = logIndex_sig)
OMsetup
```

2) Pick subset of OM settings and vary seasonality of q effect in EM
```{r}
##### EM
# EM factorial settings
miss_season <- c("BOTH", "ONE", "NONE") # Number of seasons with correctly specified effect, assumes 2 survey indices (spring & fall)
miss_q <- c("NoEnv", "NoEnvRand", "WithEnvRand", "WithEnv") # Catchability setup for EM
expand.grid(miss_season = miss_season, miss_q = miss_q)
```

3) May run sensitivity run with no ecov effect in OM but compare same EMs
```{r}
# Catchability set up
q_Ecov <- c("includeEcov", "excludeEcov") # May want to include these as a sensitivity run to compare what happens if OM has no ecov effect but EM assumes there is one (ecov process setup won't matter since fitting to data without effect specified in model)

##### OM sensitivity runs
# Factorial combinations when no ecov effect in OM
expand.grid(q_Ecov = q_Ecov, F_hist = F_hist, ageComp_sig = ageComp_sig, logIndex_sig = logIndex_sig) # probably do a subset, we have to decide what subset
```

@@

# Loop over rows in OMsetup to initialize OM and EMs
for(iom in 1:nrow(OMsetup)){
  # Pull generic input settings
  input <- genericInput

  # Ecov setup
    Ecov <- list(label = "Ecov",
                 mean = matrix(rep(0,length(groundfish_info$years)), ncol =1), # Mean = 0
                 logsigma = matrix(log(OMsetup$Ecov_process_obs_sig[iom]), length(groundfish_info$years), ncol=1),
                 years = groundfish_info$years, 
                 use_obs = matrix(rep(TRUE, length(groundfish_info$years)), ncol = 1),
                 lag = 0, 
                 where = "q", # Where/how/indices settings need to change if we do sensitivity runs
                 how = 1,
                 indices = list(1,2),
                 process_model = "ar1", # End generic inputs for ecov
                 process_mean_vals = rep(0,length(groundfish_info$years)), # Mean = 0
                 process_sig_vals = rep(OMsetup$Eco_process_sig[iom]),
                 process_cor_vals = rep(OMsetup$Ecov_process_cor[iom]),
                 beta_vals = list(rep(list(matrix(rep(OMsetup$Ecov_effect[iom], length(groundfish_info$ages)), nrow = 1)), 4))) #!!! Why 4 rows?

@@

#   # Fishing history
#   # input =
#           set_F_scenario(input, Fhist, Fmsy = Fmsy, max_mult = max_mult_Fmsy, min_mult= min_mult_Fmsy,
#     change_time = F_change_time) # Function pulled from get_FMSY.R
#   
# F_hist <- c("H-MSY", "MSY") # High-then MSY vs. MSY for entire history Fmsy is an option

# Observation error - set L-N SD parameters for catch and index age comps
input$par$catch_paa_pars[,1] <- log(OMsetup[iom, "ageComp_sig"])
input$par$index_paa_pars[,1] <- log(OMsetup[iom, "ageComp_sig"])
input$data$agg_catch_sigma[] <- OMsetup[iom,"log_catch_sig"]
input$data$agg_index_sigma[] <- OMsetup[iom,"log_index_sig"]
